Efficient and Truthful Forecasting Competitions

### 2021-01-12 paper/draft1.tex
latest version from overleaf

@@

\section{ELF}
\newcommand{\elf}{M_{\mathrm{ELF}}}
The ELF mechanism $\elf$ \citep{witkowski2018incentive}, improves upon $M_s$, by incentivizing forecasters to be truthful. For each event $k$, we hold a lottery to choose a winner. Forecaster $i$ is chosen with probability
\[ f_{i, k} = \frac{1}{n} + \frac{1}{n} \left(R_q(y_{i, k}, x_k) - \frac{\sum_{j\neq i} R_q(y_{j, k}, x_k)}{n-1} \right)\]
This score is the additively normalized quadratic score among all forecasters. A forecaster increasing their own quadratic score increases the chances they win an event, while decreasing the chances others win that event. Let $w_k$ be the winner of event $k$. Then we choose the forecaster $\elf(y, x) = \argmax_i \sum_{k=1}^m \1(w_k = i)$ as the winner of the competition. Essentially, we give the winner of each event lottery a point, and then pick the forecaster who got the most points as the overall winner.

@@

\section{SHELF}
\newcommand{\shelf}{M_{\mathrm{SHELF}}}

@@

@@

>>>>

\newcommand{\elf}{M_{\mathrm{ELF}}}
\newcommand{\shelf}{M_{\mathrm{SHELF}}}

@@

\section{ELF}
The ELF mechanism $\elf$ \citep{witkowski2018incentive}, improves upon $M_s$, by incentivizing forecasters to be truthful. For each event $k$, we hold a lottery to choose a winner. Forecaster $i$ is chosen with probability
\[ f_{i, k} = \frac{1}{n} + \frac{1}{n} \left(R_q(y_{i, k}, x_k) - \frac{\sum_{j\neq i} R_q(y_{j, k}, x_k)}{n-1} \right)\]
This score is the additively normalized quadratic score among all forecasters. A forecaster increasing their own quadratic score increases the chances they win an event, while decreasing the chances others win that event. Let $w_k$ be the winner of event $k$. Then we choose the forecaster $\elf(y, x) = \argmax_i \sum_{k=1}^m \1(w_k = i)$ as the winner of the competition. Essentially, we give the winner of each event lottery a point, and then pick the forecaster who got the most points as the overall winner.

@@

\section{SHELF}

@@

\newpage
\section{Raf notes}

Let's study the one-round version of ELF + SHELF + noise.
And actually, we won't really need to use SHELF per se.
Let's actually define
$M_{f,\sigma}(y,x) = \max_{i\in[n]} f(y,x)_i + X_i$ where the $X_i \sim N(0,\sigma^2)$ are indpendent, and $f$ is any function with $f(y,x) \in [0,m]^n$ where if you change only one entry of $y$ by $c$, then $f$ changes by at most $c^2$.
(For example, $f$ could be the sum of scores.)
Let $M_\alpha$ be the mechanism that runs $M_{f,\sigma}$ with probability $\alpha$ and $\elf$ with probability $1-\alpha$.

In what follows, when I omit $x$, I'm implicitly taking the randomness over $x$ into account (from $\theta$).
Assume $y$ is truthful for forecaster $i$ and $y'$ is any vector that differs only in $i$'s forecast on a particular round.
And actually let's focus on the case $m=1$, so there is only one round anyway.
From a direct calculation (see also a Prop.\ in \cite{witkowski2018incentive}), we have, from forecaster $i$'s point of view,
\begin{align}
  \Pr[\elf(y)=i] - \Pr[\elf(y')=i] = \tfrac 1 n (y_i - y'_i)^2~,
\end{align}
since forecaster $i$ believes $y_i = \theta_1$.

By differential privacy (I have not checked this carefully!), and the fact that $\exp(c) \leq 1+2c$ for $c\in[0,1]$, we have, for each $x$,
\begin{align}
  \Pr[M_{f,\sigma}(y',x)=i]
  & \leq \exp\left( \frac{(y_i-y'_i)^2}{2\sigma^2} \right) \Pr[M_{f,\sigma}(y,x)=i]
  \\
  & \leq \left(1 + \frac{(y_i-y'_i)^2}{\sigma^2}\right) \Pr[M_{f,\sigma}(y,x)=i]
  \\
  & \leq \Pr[M_{f,\sigma}(y,x)=i] + \frac{(y_i-y'_i)^2}{\sigma^2}~.
\end{align}

Combining, the probability of $i$ winning for $y$ and $y'$ can be compared as follows.
\begin{align}
  &\Pr[M_\alpha(y)=i]-\Pr[M_\alpha(y')=i]\\
  &=\alpha\left(\Pr[M_{f,\sigma}(y)=i] - \Pr[M_{f,\sigma}(y')=i]\right)\\
  &\qquad + (1-\alpha)\left(\Pr[\elf(y)=i] - \Pr[\elf(y')=i]\right)~.
  \\
  &\geq -\alpha \frac{(y_i-y'_i)^2}{\sigma^2} + (1-\alpha) \frac{(y_i-y'_i)^2}{n}
  \\
  &\geq \left((1-\alpha) \frac{1}{n} -\alpha \frac{1}{\sigma^2}\right) (y_i-y'_i)^2~.
\end{align}
Thus, setting $\sigma^2 = n \tfrac \alpha {1-\alpha}$ gives truthfulness, e.g.\ $\sigma^2=n$ for $\alpha=\tfrac 1 2$.
Which is maybe not what we were hoping for, but it's something.

@@

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
